Modelling BCG vaccination in the UK

### 2019-07-31 rmds/09-model-fitting.Rmd
fig caption editing

In this model fitting pipeline I have used a maximum of 5 iterations of, manual, independent proposal adaption, drawing 250 samples in each iteration, starting with the prior distributions and halfing the scale each time. The proposal distribution is also truncated so that parameter samples cannot be drawn from areas without support from the prior distributions. As for the particle calibration, I initially used a maximum likelihood method to provide a point estimate of the best fitting parameter set, followed by 1000 PMCMC steps, using a 256 particle filter. This means that the proposal distribution is adapted near to the posterior mean rather than in the tails of the posterior distribution.

@@

6. Manual independent proposal adaption at the approximate mean of the posterior distribution. The size of the proposal distribution is halved each iteration, with at most 5 iterations of adaption. The minimum target acceptance rate specified was 10% and the maximum was 20%. 250 PMCMC samples were used each time to estimate the acceptance rate. Proposal adaption was only carried out for the main model scenario, with other scenarios using this proposal.

@@

Varying the proposal size, rejuvernation thresold, and number of rejuvernation steps showed some promise at improving the quality of the model fit but ultimately computational constraints limited how much progress could be made using this approach. It is possible that a much longer runtime could result in an improved fit to data with no other model changes. Alternatively these results may driven by the model being too complex for an SMC-SMC approach to viable in this framework.

@@

```{r 09-scenarios-table, eval = FALSE}
tibble(Scenario = names(scenario_dics) %>% str_split("_") %>% 
         map_chr(~paste(., collapse = " ")), 
       DIC = unlist(transpose(scenario_dics)$dic)) %>% 
  arrange(DIC) %>% 
  mutate(Scenario = Scenario %>% 
           str_replace("plus", "and") %>% 
           str_replace("baseline", "Baseline") %>% 
           str_replace("transmission", "Transmission variable in young adults (15-29)") %>% 
           str_replace("mixing", "non-UK born mixing variable in young adults (15-29)") %>% 
  str_to_sentence()) %>% 
mutate(DIC = round(DIC, 0)) %>% 
kable(
  caption = "DIC values for each scenario evaluated during model fitting - arranged from best to worst model fit.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(latex_options = c("hold_position"))
```

@@

```{r 09-table-incidence-preds, eval = FALSE}
sum_obs_table %>% 
  set_names(., nm = c(names(.)[-ncol(.)], "Predicted Incidence (95% CI)*")) %>% 
  select(-Observation) %>% 
kable(
  caption.short = "Observed versus predicted overall TB incidence for years that the model was fitted too.", 
  caption = "Observed versus predicted overall TB incidence for years that the model was fitted too. The model underpredicted incidence for all years fitted too. The upper credible interval of the model predictions was substantially lower than the observed data. ", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(latex_options = c("hold_position")) %>% 
  add_footnote(label = "* 95% credible interval estimated using the 2.5% quantile and the 97.5% quantile.", notation = "none")
```

>>>>

In this model fitting pipeline I have used a maximum of 5 iterations of, manual, independent proposal adaption, drawing 250 samples in each iteration, starting with guassian distributions for each parameter, truncated by the range of the prior, with the mean based on the current parameter value. The standard deviation for each parameter was assumed to be the standard deviation of the prior if it was guassian and otherwise assumed to be the range of the prior if it was uniform. For each iteration I halved the  size of the standard deviation of each parameter. As for the particle calibration, I initially used a maximum likelihood method to provide a point estimate of the best fitting parameter set, followed by 1000 PMCMC steps, using a 256 particle filter. This means that the proposal distribution is adapted near to the posterior mean rather than in the tails of the posterior distribution.

@@

6. Manual independent proposal adaption at the approximate mean of the posterior distribution. It is assumed that the proposal for each parameter is guassian, truncated by the range of the prior with the mean based on the current parameter value. The standard deviation of the proposal distribution is halved each iteration, with at most 5 iterations of adaption. The minimum target acceptance rate specified was 10% and the maximum was 20%. 250 PMCMC samples were used each time to estimate the acceptance rate. Proposal adaption was only carried out for the main model scenario, with other scenarios using this proposal.

@@

Varying the proposal size, rejuvernation thresold, and number of rejuvernation steps showed some promise at improving the quality of the model fit but ultimately computational constraints limited how much progress could be made using this approach. It is possible that a much longer runtime could result in an improved fit to data with no other model changes. Alternatively these results may driven by the model being too complex for an SMC-SMC approach to be viable in this framework.

@@

```{r 09-scenarios-table}
tibble(Scenario = names(scenario_dics) %>% str_split("_") %>% 
         map_chr(~paste(., collapse = " ")), 
       DIC = unlist(transpose(scenario_dics)$dic)) %>% 
  arrange(DIC) %>% 
  mutate(Scenario = Scenario %>% 
           str_replace("plus", "and") %>% 
           str_replace("baseline", "Baseline") %>% 
           str_replace("transmission", "Transmission variable in young adults (15-29)") %>% 
           str_replace("mixing", "non-UK born mixing variable in young adults (15-29)") %>% 
  str_to_sentence()) %>% 
mutate(DIC = round(DIC, 0)) %>% 
kable(
  caption = "DIC values for each scenario evaluated during model fitting - arranged from best to worst model fit.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(latex_options = c("hold_position"))
```

@@

```{r 09-table-incidence-preds}
sum_obs_table %>% 
  set_names(., nm = c(names(.)[-ncol(.)], "Predicted Incidence (95% CI)*")) %>% 
  select(-Observation) %>% 
kable(
  caption.short = "Observed versus predicted overall TB incidence for years that the model was fitted too.", 
  caption = "Observed versus predicted overall TB incidence for years that the model was fitted too. The model underpredicted incidence for all years fitted too. The upper credible interval of the model predictions was substantially lower than the observed data. ", booktabs = TRUE, longtable = TRUE, escape = TRUE) %>% 
  kable_styling(latex_options = c("hold_position")) %>% 
  add_footnote(label = "* 95% credible interval estimated using the 2.5% quantile and the 97.5% quantile.", 
               notation = "none")
```

### 2019-07-31 rmds/10-model-results.Rmd
fig caption editing

```{r 10-table-future-incidence, eval = FALSE}
sum_scenario %>% filter(Year %% 5 == 0) %>% 
  set_names(., nm = names(.) %>% str_replace_all("BCG", "BCG (95% CI)\\*")) %>% 
  select(-Observation) %>% 
kable(
  caption.short = "", 
  caption = "", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(latex_options = c("hold_position")) %>% 
  add_footnote(label = "* 95% credible interval estimated using the 2.5% quantile and the 97.5% quantile.", notation = "none")
```

@@

```{r 10-table-deaths-preds, eval = FALSE}
sum_deaths_scenario %>% filter(Year %% 5 == 0) %>% 
  set_names(., nm = names(.) %>% str_replace_all("BCG", "BCG (95% CI)\\*")) %>% 
  select(-Observation) %>% 
kable(
  caption.short = "", 
  caption = "", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(latex_options = c("hold_position")) %>% 
  add_footnote(label = "* 95% credible interval estimated using the 2.5% quantile and the 97.5% quantile.", notation = "none")
```

>>>>

```{r 10-table-future-incidence}
sum_scenario %>% filter(Year %% 5 == 0) %>% 
  set_names(., nm = names(.) %>% str_replace_all("BCG", "BCG (95% CI)\\*")) %>% 
  select(-Observation) %>% 
kable(
  caption.short = "", 
  caption = "", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(latex_options = c("hold_position")) %>% 
  add_footnote(label = "* 95% credible interval estimated using the 2.5% quantile and the 97.5% quantile.", notation = "none")
```

@@

```{r 10-table-deaths-preds}
sum_deaths_scenario %>% filter(Year %% 5 == 0) %>% 
  set_names(., nm = names(.) %>% str_replace_all("BCG", "BCG (95% CI)\\*")) %>% 
  select(-Observation) %>% 
kable(
  caption.short = "", 
  caption = "", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(latex_options = c("hold_position")) %>% 
  add_footnote(label = "* 95% credible interval estimated using the 2.5% quantile and the 97.5% quantile.", notation = "none")
```
